\documentclass[12pt]{article}

% TEMPLATE DEFAULT PACKAGES
\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,graphicx,color,setspace,sectsty,comment,natbib,pdflscape,array,adjustbox}

% ADDED PACKAGES FOR THIS MANUSCRIPT
\usepackage{palatino,newtxmath,multirow,titlesec,threeparttable,tabu,booktabs,titlesec,threeparttable,mathtools,bm,bbm,subcaption,pdflscape,tcolorbox,mathrsfs,tikz,float,longtable}
% endfloat,
% \usepackage[capposition=top]{floatrow}
\usepackage{tikzpagenodes}
\usepackage{pbox}

\usepackage{afterpage}
\usepackage[hyphens]{url}
\usepackage[margin=1cm]{caption}

\usepackage[draft]{hyperref}
\newcommand{\tim}{$\,\times\,$}
% FIGURES & TABLES CAPTION STYLING
\captionsetup[figure]{labelfont={bf},name={Figure},labelsep=period}
\captionsetup[table]{labelfont={bf},name={Table},labelsep=period}

% SECTION TITLE SETTINGS
\titlelabel{\thetitle.\enskip}
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

% COLUMN TYPES
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}{>{\centering\arraybackslash}p{5.2em}}
\newcolumntype{D}{>{\centering\arraybackslash}p{5em}}
\newcolumntype{G}{>{\centering\arraybackslash}p{6em}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}


% MARGINS AND SPACING
\normalem
\geometry{left=1.1in,right=1.1in,top=1.0in,bottom=1.0in}
\setlength{\parskip}{2.5pt}

% SPECIAL CELL 
\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

% NO INDENT ON FOOTNOTES
\usepackage[hang,flushmargin]{footmisc}


\newcommand{\regtextfirst}{
	``hm'' refers to a hectometer or 100 meters. Observations are one hectare plots ($\text{hm}^{2}$).
``Plot footprint'' is the effect of an additional hectare of overlap with constructed projects.  
For plots outside of projects, ``Plot neighborhood'' is the effect of an additional hectare of overlap between 0 to 5 hm rings around plots and constructed projects.  Specifications also include separate effects for 5 hm rings from 0 to 40 hm.
Standard errors are clustered at the project level (in parentheses). 
\textsuperscript{c} p$<$0.10,\textsuperscript{b} p$<$0.05,\textsuperscript{a} p$<$0.01 \,\,
}

\newcommand{\regtext}{
	``hm'' refers to a hectometer or 100 meters. Observations are one hectare plots ($\text{hm}^{2}$).
``Plot footprint'' is the effect of an additional hectare of overlap with constructed projects.  
For plots outside of projects, ``Plot neighborhood'' is the effect of an additional hectare of overlap between 0 to 5 hm rings around plots and constructed projects.  Although not displayed in the table, specifications also include separate effects for 5 hm rings from 0 to 40 hm and results are available upon request.
Standard errors are clustered at the project level (in parentheses). 
\textsuperscript{c} p$<$0.10,\textsuperscript{b} p$<$0.05,\textsuperscript{a} p$<$0.01 \,\,
}

\newcommand{\ballparktext}{
	``Footprint'' sums the footprint effect across plots within constructed projects.
	``Spillover'' sums the neighborhood effect (0-.5 km) across plots nearby constructed projects.
	``Total'' sums ``Footprint'' and ``Spillover'' effects.
}

\newcommand{\hmref}{
	``hm'' refers to a hectometer or 100 meters.
}
\newcommand{\haref}{
	``ha'' refers to a hectare or one $\text{hm}^{2}$.
}

\newcommand{\hmrefha}{
	``hm'' refers to a hectometer or 100 meters. ``ha'' refers to a hectare or one $\text{hm}^{2}$.
}




\begin{document}

\begin{table}
\small
\centering
\caption{Effects on House Quality (2001-2011)}\label{table:inf_census}
\vspace{-2mm}
\resizebox{1\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}{lCCCCC}
\toprule
\input{figures/inf_census_top}
\bottomrule
\end{tabular}
\begin{tablenotes}
\item 
\end{tablenotes}
\end{threeparttable}
}
\end{table}

\begin{table}
\small
\centering
\caption{Effects on House Quality Placebo (1996-2001)}\label{table:inf_census_placebo}
\vspace{-2mm}
\resizebox{1\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}{lCCCCC}
\toprule
\input{figures/inf_census_test_top}
\bottomrule
\end{tabular}
\begin{tablenotes}
\item 
\end{tablenotes}
\end{threeparttable}
}
\end{table}



\begin{table}
\small
\centering
\caption{Effects on Census Outcomes (2001-2011)}\label{table:demo}
\vspace{-2mm}
\resizebox{1\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}{lCCCCC}
\toprule
\input{demo_true_test_top}
\bottomrule
\end{tabular}
\begin{tablenotes}
\item 
\end{tablenotes}
\end{threeparttable}
}
\end{table}



\begin{table}
\small
\centering
\caption{Effects on Census Outcomes Placebo (1996-2001)}\label{table:demo_placebo}
\vspace{-2mm}
\resizebox{1\linewidth}{!}{
\begin{threeparttable}
\begin{tabular}{lCCCCC}
\toprule
\input{figures/demo_test_top}
\bottomrule
\end{tabular}
\begin{tablenotes}
\item 
\end{tablenotes}
\end{threeparttable}
}
\end{table}

\end{document}


